Specialisation in food webs scales with species richness but not with latitude

### 2015-06-29 manuscript/latitudepaper.tex
Updated text

\title{Generality in food webs scales with species richness but not with latitude}
\author{Alyssa R. Cirtwill$^{1,2}$, Daniel B. Stouffer$^{1,3}$, Tamara N. Romanuk$^{2}$}
\date{\small$^1$School of Biological Sciences\\University of Canterbury\\
Private Bag 4800\\Christchurch 8140, New Zealand \\
\medskip$^2$Department of Biology\\
Life Science Centre, Dalhousie University\\1355 Oxford St., P0 BOX 15000\\
Halifax NS, B3H 4R2, Canada\\
\medskip$^3$Centre for Integrative Ecology\\School of Biological Sciences\\University of Canterbury\\
Private Bag 4800\\Christchurch 8140, New Zealand \\}

@@

  Link density, generality, and vulnerability each increased with increasing
  species richness ($\beta_0$=0.644, $p$\textless0.001; $\beta_0$=0.557,
  $p$\textless0.001; and $\beta_0$=0.644, $p$\textless0.001, respectively;
  Fig~\ref{props_v_lat}).For estuarine, marine, and terrestrial food webs the
  strength of this scaling did not vary with latitude
  ($\beta_{Latitude}$=-0.001, $p$=0.206 for link density;
  $\beta_{Latitude}$=-0.001, $p$=0.373 for generality; and
  $\beta_{Latitude}$=-0.001, $p$=0.207 for vulnerability; Fig.~\ref{S}). In
  lake food webs, however, the scaling of each property was stronger towards
  the poles ($\beta_{Latitude:Lake}$=0.005, $p$=0.005;
  $\beta_{Latitude:Lake}$=0.006, $p$\textless0.001; and
  $\beta_{Latitude:Lake}$=0.005, $p$=0.005, respectively). In stream food
  webs, generality increased more rapidly towards the poles
  ($\beta_{Latitude:Stream}$=0.007, $p$\textless0.001) while link density and
  vulnerability did not vary with latitude.

  Unlike species richness, only generality showed an overall increase with increasing
  proportions of basal resources in a web ($\beta_0$=-0.030, $p$=0.801;
  $\beta_0$=0.328, $p$=0.013; and $\beta_0$=-0.030, $p$=0.802 for link
  density, generality, and vulnerability, respectively;
  Fig.~\ref{props_v_lat}). In each case, however, there were substantial
  effects of latitude on the scaling relationships. Thus, the scaling
  exponents for link density and vulnerability were negative for estuarine,
  lake, and stream communities and became more negative towards the poles
  ($\beta_{Latitude}$=-0.007, $p$=0.006 and $\beta_{Latitude}$=-0.007,
  $p$=0.005, respectively; Fig.~\ref{BIT}). For the scaling of link density and
  vulnerability in marine and terrestrial communities, the overall effect of
  latitude on scaling was positive ($\beta_{Latitude:Marine}$=0.009,
  $p$=0.038; $\beta_{Latitude:Terrestrial}$=0.012, $p$=0.003 and
  $\beta_{Latitude:Marine}$=0.009, $p$=0.037;
  $\beta_{Latitude:Terrestrial}$=0.012, $p$=0.003, respectively) such that
  there was little overall scaling relationship. In contrast, the scaling
  exponent of generality with the proportion of basal resources was positive
  near the equator and negative near the poles in estuarine, lake, and stream
  communities ($\beta_{Latitude}$=-0.009, $p$=0.002) while the scaling
  exponents for marine and terrestrial communities were not significantly
  different from zero ($\beta_{Latitude:Marine}$=0.011, $p$=0.021;
  $\beta_{Latitude:Terrestrial}$=0.014, $p$=0.002.

  As with species richness, link density, generality, and vulnerability
  generally increased with the proportion of intermediate consumers in a web
  ($\beta_0$=2.10, $p$\textless0.001; $\beta_0$=0.592, $p$=0.006; and
  $\beta_0$=0.843, $p$\textless0.001 respectively; Fig.~\ref{props_v_lat}). 
  The best-fit models for link density and vulnerability did not include
  any effect of latitude. In contrast, the strength of the scaling relationship
  for generality increased weakly towards the poles in estuarine, marine, and
  terrestrial food webs ($\beta_{Latitude}$=0.003, $p$=0.486) but decreased 
  towards the poles in lakes and streams ($\beta_{Latitude:Lake}$=-0.019, $p$=0.027
  and $\beta_{Latitude:Stream}$=-0.031, $p$=0.002; Fig~\ref{BIT}).

  In contrast to the generally positive scaling relationships above, link
  density, generality, and vulnerability decreased as the proportion of top
  predators increased ($\beta_0$=-0.584, $p$\textless0.001; $\beta_0$=-0.418,
  $p$\textless0.001; and $\beta_0$=-0.585, $p$\textless0.001 respectively;
  Fig~\ref{props_v_lat}). For estuarine, lake, and marine food webs the
  strength of this scaling did not vary significantly with latitude
  ($\beta_{Latitude}$=0.001, $p$=0.652 for link density;
  $\beta_{Latitude}$=0.001, $p$=0.654 for vulnerability) except in stream food
  webs, where scaling was more strongly negative in the tropics and near zero
  near the poles ($\beta_{Latitude:Stream}$=0.010, $p$=0.006 for link density;
  $\beta_{Latitude:Stream}$=0.010, $p$=0.007 for vulnerability;
  Fig~\ref{BIT}). The best-fit model for generality did not include any terms
  for latitude in any ecosystem type). In stream food webs, scaling was more
  strongly negative in the tropics and did not differ from zero near the poles

>>>>

\title{Specialisation in food webs scales with species richness but not with latitude}
\author{Alyssa R. Cirtwill$^{1,2,3}$, Daniel B. Stouffer$^{1,2}$, Tamara N. Romanuk$^{3}$}
\date{\small$^1$School of Biological Sciences\\University of Canterbury\\
Private Bag 4800\\Christchurch 8140, New Zealand \\
\medskip$^2$Centre for Integrative Ecology\\School of Biological Sciences\\University of Canterbury\\
Private Bag 4800\\Christchurch 8140, New Zealand \\
\medskip$^3$Department of Biology\\
Life Science Centre, Dalhousie University\\1355 Oxford St., P0 BOX 15000\\
Halifax NS, B3H 4R2, Canada\\}

@@

  Link density, generality, and vulnerability each increased with increasing
  species richness ($\beta_0$=0.637, $p$\textless0.001; $\beta_0$=0.553,
  $p$\textless0.001; and $\beta_0$=0.637, $p$\textless0.001, respectively;
  Fig~\ref{props_v_lat}).For estuarine, marine, and terrestrial food webs the
  strength of this scaling did not vary with latitude
  ($\beta_{Latitude}$=-0.001, $p$=0.365 for link density;
  $\beta_{Latitude}$=-0.001, $p$=0.535 for generality; and
  $\beta_{Latitude}$=-0.001, $p$=0.366 for vulnerability; Fig.~\ref{S}). In
  lake food webs, however, the scaling of each property was stronger towards
  the poles ($\beta_{Latitude:Lake}$=0.004, $p$=0.019;
  $\beta_{Latitude:Lake}$=0.005, $p$=0.004; and
  $\beta_{Latitude:Lake}$=0.004, $p$=0.018, respectively). In stream food
  webs, generality increased more rapidly towards the poles
  ($\beta_{Latitude:Stream}$=0.007, $p$=0.001) while link density and
  vulnerability did not vary with latitude.

  Unlike species richness, only generality showed an overall increase with
  increasing proportions of basal resources in a web ($\beta_0$=0.019,
  $p$=0.859; $\beta_0$=0.383, $p$=0.001; and $\beta_0$=0.019, $p$=0.855 for
  link density, generality, and vulnerability, respectively;
  Fig.~\ref{props_v_lat}). In each case, however, there were substantial
  effects of latitude on the scaling relationships. Thus, the scaling
  exponents for link density and vulnerability were negative for estuarine,
  lake, and stream communities and became more negative towards the poles
  ($\beta_{Latitude}$=-0.008, $p$\textless0.001 and $\beta_{Latitude}$=-0.008,
  $p$\textless0.001, respectively; Fig.~\ref{BIT}). For the scaling of link
  density and vulnerability in marine and terrestrial communities, the overall
  effect of latitude on scaling was positive ($\beta_{Latitude:Marine}$=0.011,
  $p$=0.020; $\beta_{Latitude:Terrestrial}$=0.016, $p$\textless0.001 and
  $\beta_{Latitude:Marine}$=0.011, $p$=0.020;
  $\beta_{Latitude:Terrestrial}$=0.016, $p$\textless0.001, respectively) such
  that there was little overall scaling relationship. In contrast, the scaling
  exponent of generality with the proportion of basal resources was positive
  near the equator and negative near the poles in estuarine, lake, and stream
  communities ($\beta_{Latitude}$=-0.010, $p$\textless0.001) while there was
  very little change in the scaling exponents for marine and terrestrial
  communities ($\beta_{Latitude:Marine}$=0.013, $p$=0.010;
  $\beta_{Latitude:Terrestrial}$=0.019, $p$\textless0.001).

  As with species richness, link density, generality, and vulnerability
  generally increased with the proportion of intermediate consumers in a web
  ($\beta_0$=0.771, $p$\textless0.001; $\beta_0$=0.349, $p$=0.235; and
  $\beta_0$=0.771, $p$\textless0.001 respectively; Fig.~\ref{props_v_lat}). 
  The best-fit models for link density and vulnerability did not include
  any effect of latitude. In contrast, the strength of the scaling relationship
  for generality increased weakly towards the poles in estuarine, marine, and
  terrestrial food webs ($\beta_{Latitude}$=0.008, $p$=0.204) but decreased 
  towards the poles in lakes and streams ($\beta_{Latitude:Lake}$=-0.021, $p$=0.018
  and $\beta_{Latitude:Stream}$=-0.035, $p$=0.001; Fig~\ref{BIT}).

  In contrast to the generally positive scaling relationships above, link
  density, generality, and vulnerability decreased as the proportion of top
  predators increased ($\beta_0$=-0.532, $p$\textless0.001; $\beta_0$=-0.454,
  $p$\textless0.001; and $\beta_0$=-0.532, $p$\textless0.001 respectively;
  Fig~\ref{props_v_lat}). For estuarine, lake, and marine food webs the
  strength of this scaling did not vary significantly with latitude
  ($\beta_{Latitude}$=-0.001, $p$=0.481 for link density;
  $\beta_{Latitude}$=-0.001, $p$=0.472 for vulnerability) except in stream and terrestrial food
  webs, where scaling was more strongly negative in the tropics and near zero
  near the poles ($\beta_{Latitude:Stream}$=0.012, $p$=0.001 and $\beta_{Latitude:Terrestrial}$=0.009, $p$=0.015 for link density;
  $\beta_{Latitude:Stream}$=0.012, $p$=0.001 and
  $\beta_{Latitude:Terrestrial}$=0.009, $p$=0.015
  for vulnerability;
  Fig~\ref{BIT}). The best-fit model for generality did not include any terms
  for latitude in any ecosystem type.
